Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
improving tensor description

\begin{example}
For a $2$-dimensional situation, 
\begin{align*}
& \interp{\stenCenS{2}{1}}_2 \\
= \; & \{i\singleEntry{1}{2} \,|\, i \in \{\text{-}2,..., 2\}\!\setminus\!0 \}
  \cup  \{ \zeroEntry{1}{2} \,|\, \textit{true} \} \\
= \; & \{-2\singleEntry{1}{2}, \, -\singleEntry{1}{2}, \,
  \singleEntry{1}{2}, \, 2\singleEntry{1}{2}\} \cup
  \{\zeroEntry{1}{2}\} \\
= \; & \{\vtwoh{-2}{\infty}, \vtwoh{-1}{\infty}, \vtwoh{0}{\infty}, 
\vtwoh{1}{\infty}, \vtwoh{2}{\infty}\} 
\end{align*}
Note the absboring behaviour of $\infty$ with respect to
multiplication. 
\end{example}
\noindent
The model of $\term{+}$ is the straightforward union of 
the models of subterms. The model of $\term{*}$ is uses
an intermediate tensor $\otimes$ which requires some intermediate
definitions.

\begin{definition}[Constrained dimensions]
Given an $n$-dimensional model $M$, the \emph{constrained dimensions}
of that model is a subset of $\{1, \ldots, n\}$ giving
the dimensions for which at least one scheme $u \in M$ has
a non-$\infty$ value in that dimension. That is:
%
\begin{align*}
\effdims{M}{n} =
\bigcup_{u \in M} \{i \mid i \in \{1,\ldots,n\} \, \wedge \, u_i \neq
  \infty\}
\end{align*}
For example, $\effdims{\{\vtwoh{0}{\infty},\vtwoh{\infty}{\infty}\}}{2}
 = \{1\}$.
\end{definition}

\begin{definition}The tensor $\otimes_n$ of $n$-dimensional
models is defined:
%
%\begin{align*}
%\{\vtwoh{0}{0}, \vtwoh{1,1}\}
% \otimes \{\vtwoh{0,\infty}\}
%\end{align*}
%
%
%%
\begin{alignat*}{2}
& M \otimes_n N =
  \Bigg\{x \; \Bigg| \;
    \parbox{6cm}{$(u, v) \in M \times N $ \\[0.1em]
                 $\;\; \wedge \;\, j \in (\effdims{M}{n} \cup
                   \effdims{N}{n})$\\[0.1em]
                  $\;\; \wedge \;\, i \in \{1, \ldots, 2^n\} \, \wedge
                  \, x = (u \bowtie v)_i$ \\[0.1em]                   
                   $\;\; \wedge \;\, x_j \neq \infty$
                  } \Bigg\} 
\end{alignat*}
The intuition for the tensor is that it gives all possible
ways of combining indexing schemes in two models
such that for every scheme
$u \in (M \otimes_n N)$ then $w \sim u$ for some $w$
\emph{iff} $\;\exists v \in M . w \sim v$ and
$\exists v' \in N . w \sim v'$. Thus it acts as a kind of intersection
for indexing schemes. We unpack its definition in pieces.

We start with the first guard taking the Cartesian product of
the two models and quantifying over all pairings $(u, v)$. 
The second guard says that $j$ is in the set of constrainted
dimensions for $M$ and the constrained dimensions for $N$.
That is, every $j$ corresponds to a dimension in which 
either $M$ or $N$ has a scheme $u$ where $u_j \neq \infty$,
and thus is constraining.

%for all possible
%pairs of schemes $(u, v)$ in the Cartesian product of the
%two models $M \times N$, then

% combines non-$\infty$ constraints
%of the unifiers in different dimensions into less accepting unifiers.
%This means an access pattern unifying with this unifier would require
%both models to be satisfied. Use of $dims$ gives the dimensions that
%constitutent models talk about and is used in the product to eliminate
%general unifier appearing in these dimensions. Allowing general unifier
%in these dimensions would prevent both models to be satisfied
%simultaneously.

The $\bowtie$ operation, which we call \emph{pairwise permutation}
(described below), constructs a $2^n \times n$ matrix ($2^n$-vector of
$n$-vectors), which is indexed by $i$ above
and then $j$. Using this, the $\otimes$ operation constructs the set
of schemes which are the 
pairwise permutations of every pair $(u, v)$ of schemes in the Cartesian
product of $M \times N$. These permuted schemes are only those
which have non-$\infty$ components in the constrained dimensions
given by each model.

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by
non-deterministic picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\setlength{\arraycolsep}{0.5em}
\left[\begin{array}{ccc}
0 & 1 & 2 \\
0 & 1 & 5 \\
0 & 4 & 2 \\
0 & 4 & 5 \\
3 & 1 & 2 \\
3 & 1 & 5 \\
3 & 4 & 2 \\
3 & 4 & 5
\end{array}
\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. Thus, we can
defined the pairwise permutation matrix as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
where $b$ is a logical matrix where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$.
%%
\todo{Complete this}
\end{definition}

>>>>

\begin{example}
For a $2$-dimensional situation, 
\begin{align*}
& \interp{\stenCenS{2}{1}}_2 \\
= \; & \{i\singleEntry{1}{2} \,|\, i \in \{\text{-}2,..., 2\}\!\setminus\!0 \}
  \cup  \{ \zeroEntry{1}{2} \,|\, \textit{true} \} \\
= \; & \{-2\singleEntry{1}{2}, \, -\singleEntry{1}{2}, \,
  \singleEntry{1}{2}, \, 2\singleEntry{1}{2}\} \cup
  \{\zeroEntry{1}{2}\} \\
= \; & \{\vtwoh{-2}{\infty}, \vtwoh{-1}{\infty}, \vtwoh{0}{\infty}, 
\vtwoh{1}{\infty}, \vtwoh{2}{\infty}\} 
\end{align*}
Note the absboring behaviour of $\infty$ with respect to
multiplication. 
\end{example}
\noindent
The model of $\term{+}$ is the straightforward union of 
the models of subterms. The model of $\term{*}$ is
the tensor $\otimes$ of subterm models. The tensor 
is non-trivial; we give some informal intutition first.

\paragraph{$\otimes$ of models}

The intuition behind $\otimes$ is that it takes
all possible pairs of indexing schemes, treating them as the
lower and upper bounds of an $n$-dimensional rectangle from which the 
remaining vertices of the rectangle are generated (like a bounding
box). \eg{}:
%%
\begin{align*}
\{\vtwoh{1}{2}, \vtwoh{5}{6}\} \otimes \{\vtwoh{3}{4}\} = 
\{&\vtwoh{1}{2}, \vtwoh{3}{2}, \vtwoh{1}{4}, \vtwoh{3}{4}, \\
  & \vtwoh{3}{6}, \vtwoh{5}{4}, \vtwoh{5}{6}\}
\end{align*}
%%
Internally $\otimes$ takes a Cartesian product of the indexing
schemes (all pairs) and then combines each pair using the 
\emph{pairwise permutation} written $\bowtie$ (defined later)
which generates all possible interleavings of two vectors
(\ie{}, all possible non-deterministic choices between $u_i$ and
$v_i$ in each dimension). Some care is then taken over
how this interacts with $\infty$: pariwse permtuation
corresponds to non-determinstic choice, in each dimension $i$,
between two values $u_i$ and $v_i$, but if either 
is $\infty$ then the resultying pairwise permutation is 
filtered out if the dimension $i$ has \emph{any other scheme
in the model where $u_i \neq \infty$}, \eg{}:
\begin{align*}
\{\vtwoh{1}{\infty}\} \otimes \{\vtwoh{3}{4}\} = 
\{\vtwoh{1}{4}, \vtwoh{3}{4}\}
\end{align*}
% 
where $\vtwoh{1}{\infty}$ and $\vtwoh{3}{\infty}$ have been filtered.
With respect to the notion of unification, the 
%
%
% combines non-$\infty$ constraints
%of the unifiers in different dimensions into less accepting unifiers.
%This means an access pattern unifying with this unifier would require
%both models to be satisfied. Use of $dims$ gives the dimensions that
%constitutent models talk about and is used in the product to eliminate
%general unifier appearing in these dimensions. Allowing general unifier
%in these dimensions would prevent both models to be satisfied
%simultaneously.
%
The notion of which dimensioons $i$ have that \emph{any scheme in the model
has $u_i \neq \infty$} is represented by the \emph{constrained
dimensions} intermediate function:
%
\begin{definition}[Constrained dimensions]
Given an $n$-dimensional model $M$, the \emph{constrained dimensions}
of that model is a subset of $\{1, \ldots, n\}$ of 
the dimensions for which at least one scheme $u \in M$ has
a non-$\infty$ value in that dimension. That is:
%
\begin{align*}
\effdims{M}{n} =
\bigcup_{u \in M} \{i \mid i \in \{1,\ldots,n\} \, \wedge \, u_i \neq
  \infty\}
\end{align*}
For example, $\effdims{\{\vtwoh{0}{\infty},\vtwoh{\infty}{\infty}\}}{2}
 = \{1\}$.
\end{definition}

\begin{definition}The tensor $\otimes_n$ of $n$-dimensional
models is defined:
\begin{alignat*}{2}
& M \otimes_n N =
  \Bigg\{x \; \Bigg| \;
    \parbox{6cm}{$(u, v) \in M \times N $ \\[0.1em]
                 $\;\; \wedge \;\, j \in (\effdims{M}{n} \cup
                   \effdims{N}{n})$\\[0.1em]
                  $\;\; \wedge \;\, i \in \{1, \ldots, 2^n\} \, \wedge
                  \, x = (u \bowtie v)_i$ \\[0.1em]                   
                   $\;\; \wedge \;\, x_j \neq \infty$
                  } \Bigg\} 
\end{alignat*}
The first guard takes the Cartesian product of
the two models and quantifies over all pairings $(u, v)$. 
The second guard says that $j$ ranges of the constrained
dimensions for $M$ unioned with the constrained dimensions for $N$.
That is, every $j$ corresponds to a dimension in which 
either $M$ or $N$ has a scheme $u$ where $u_j \neq \infty$,
and thus is constrained.

The $\bowtie$ \emph{pairwise permutation} 
constructs a $2^n \times n$ matrix ($2^n$-vector of
$n$-vectors), which is indexed by $i$ above
and then $j$. Using this, the $\otimes$ operation constructs the set
of schemes which are the 
pairwise permutations of every pair $(u, v)$ of schemes in the Cartesian
product of $M \times N$. These permuted schemes are only those
which have non-$\infty$ components in the constrained dimensions
given by each model, given by the guard $x_j \neq \infty$.

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by
non-deterministic picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\setlength{\arraycolsep}{0.5em}
\left[\begin{array}{ccc}
0 & 1 & 2 \\
0 & 1 & 5 \\
0 & 4 & 2 \\
0 & 4 & 5 \\
3 & 1 & 2 \\
3 & 1 & 5 \\
3 & 4 & 2 \\
3 & 4 & 5
\end{array}
\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. Thus, we can
defined the pairwise permutation matrix as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
where $b$ is a logical matrix where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$.
%%
\end{definition}
